Joint maximum-likelihood of phylogenies and ancestral states is not consistent

### 2018-02-06 joint_inf.tex
final edit

Previous work in phylogenetics has developed consistency counterexamples using the same four-taxon topologies used here \citep{Felsenstein1978-rr}.
In this previous work, when simulating data under the ``Felsenstein zone'' topology $\tau_2$, as the number of observations increases, the ``Farris zone'' topology $\tau_1$ becomes more likely when performing a particular estimation procedure.
This is the converse of what we have shown for joint inference, where the Felsenstein topology is more likely than the Farris topology.
Moreover, the inconsistency in \citet{Felsenstein1978-rr} is attributed to long branch attraction, i.e., the fact that there may be multiple long branches where parallel changes are more likely than a single change along a short branch.
This is not the case here; for our case, the inconsistency generally occurs when all branches are long, and has more to do with the choices of the form of the likelihood than from the interplay of long and short branches.
Difficulties in phylogenetic estimation when generating data on the ``Farris zone'' tree have been found by \citet{Siddall1998-hq}, though \citet{Swofford2001-hr} show that sequence length plays a major role in these issues.

>>>>

Previous work in phylogenetics has developed consistency counterexamples using the same four-taxon topologies used here \citep{Felsenstein1978-rr}.
In this previous work, when simulating data under the ``Felsenstein zone'' topology $\tau_2$, as the number of observations increases, the ``Farris zone'' topology $\tau_1$ becomes more likely when performing a particular estimation procedure.
This is the converse of what we have shown for joint inference, where the Felsenstein topology is more likely than the Farris topology.
Moreover, the inconsistency demonstrated by \citet{Felsenstein1978-rr} is attributed to long branch attraction, i.e., the fact that there may be multiple long branches where parallel changes are more likely than a single change along a short branch.
This is not the case here; for our case, the inconsistency generally occurs when all branches are long, and has more to do with the choices of the form of the likelihood than from the interplay of long and short branches.
Difficulties in phylogenetic estimation when generating data on the ``Farris zone'' tree have been found by \citet{Siddall1998-hq}, though \citet{Swofford2001-hr} show that sequence length plays a major role in these issues.
